---
title: "`dplyr` & `magrittr` LAB SOLUTIONS"
author: "Paul Intrevado"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include = F}
library(formatR)
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r libraries, include = F}
library(tidyverse)
library(hflights)
library(magrittr)
library(knitr)
myflights <- hflights
pokemon <- read_csv("~/Desktop/pokemon.csv")
unc <- read_csv("~/Desktop/uncSalaries.csv")
```

## Question 1

The `hflights` package contains a dataset named `hflights`, which provides information on `r format(nrow(myflights), big.mark = ",")` flights in 2011 leaving from Houston-based airports. Answer the following questions to help you practice your `dplyr` and `magrittr` skills.

\ 

- How many flights departed per month in total? From IAH per month? From HOU per month?

    ```{r}
    myflights %>% 
      count(Month, Origin) %>% 
        knitr::kable()
    ```
\ 

- What was the airline with the most total departures from IAH? From HOU?

    ```{r}
    myflights %>% 
      group_by(Origin) %>% 
        count(UniqueCarrier, sort = T) %>% 
            top_n(1) %>% 
              knitr::kable()
    ```
\ 

- How many flights were cancelled in 2011? 

    ```{r}
    myflights %>% 
        summarize(sum(Cancelled))
    ```
\ 

- Which airline suffered from the most cancelled flights?
    ```{r}
    myflights %>% 
      group_by(UniqueCarrier) %>% 
        summarise(x = sum(Cancelled)) %>% 
          arrange(desc(x)) %>% 
            slice(1)
    ```
\ 

- Which airline cancelled the most flights *relative* to their total number of flights?

    ```{r}
    myflights %>% 
      group_by(UniqueCarrier) %>% 
        summarise(x = sum(Cancelled), y = n()) %>% 
          mutate(prctCancelled = x / y * 100) %>% 
            arrange(desc(prctCancelled)) %>% 
              top_n(1)
    ```
\ 

- What are the top 3 airlines with the longest mean departure delay? 
    ```{r}
    myflights %>% 
      group_by(UniqueCarrier) %>% 
        summarize(meanDepDelay = mean(DepDelay, na.rm = T)) %>% 
          arrange(desc(meanDepDelay)) %>% 
            top_n(3) %>% 
              knitr::kable()
    ```
\newpage


- Create a table of all airlines describing the mean, median and variance of departure delay, ordered alphabetically by airline? (do this in a single pipe)
    ```{r}
    myflights %>% 
      select(UniqueCarrier, DepDelay) %>% 
        group_by(UniqueCarrier) %>% 
          summarise_all(funs(mean(., na.rm = T), median(., na.rm = T), var(., na.rm = T))) %>% 
            arrange(UniqueCarrier) %>% 
              knitr::kable()
    ```
\ 

- Which airline had the longest mean arrival delay? 
    ```{r}
    myflights %>% 
        group_by(UniqueCarrier) %>% 
          summarise(meanDepDelay = mean(ArrDelay, na.rm = T)) %>% 
            arrange(desc(meanDepDelay)) %>% 
              slice(1)
    ```
\ 

- Which on which day of the week are there the most flights?
    ```{r}
    myflights %>% 
      count(DayOfWeek, sort = T)
    ```
\ 

- Which carrier has the worst `AirTime` to `Actual Elapsed Time` ratio (the latter of which includes taxiing
    ```{r}
    myflights %>% 
      mutate(x = AirTime / ActualElapsedTime) %>% 
        group_by(UniqueCarrier) %>% 
          summarise(y = mean(x, na.rm = T)) %>% 
            arrange(y) %>% 
              slice(1)
    ```
\ 

- Which flights had a delayed departure but arrived before scheduled time?
    ```{r}
    hflights %>% 
      filter(DepDelay > 0, ArrDelay < 0) %>% 
        glimpse()
    ```
\ 

- Create a new hflights1 dataframe with an additional variable delay_percent to the dataset.
    
    ```{r}
    hflights1 <-hflights %>% 
      mutate(delay_percent = (ArrDelay - DepDelay)/DepDelay * 100)
    ```
  \ 
  
- Use `airlines` to rename the carriers 
    
    ```{r}
    airlines <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
             "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
             "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
             "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")
    ```

    ```{r}
    hflights$UniqueCarrier <- airlines[hflights$UniqueCarrier]
    ```
\ 

- Find the flights flown by one of JetBlue, American_Eagle, or Continental

    ```{r}
    hflights %>% 
      filter(UniqueCarrier %in% c('JetBlue', 'American_Eagle', 'Continental')) %>% 
        glimpse()
    ```
\ 
- Which flights had taxiing time that was greater than flying time? (where taxiing: `TaxinIn` + `TaxiOut`)
    ```{r}
    hflights %>% 
        filter((TaxiIn + TaxiOut) > AirTime) %>% 
          glimpse()
    ```
\ 

- Find all the flights that were cancelled after being delayed

    ```{r}
    hflights %>% 
      filter(DepDelay > 0, Cancelled == 1) %>% 
        glimpse()
    ```
\ 

- Display all the flights leaving IAH before 10 am and arrange according to decreasing `AirTime` 

    ```{r}
    hflights %>% 
      filter(Origin == 'IAH', DepTime < 800) %>% 
        arrange(desc(AirTime)) %>% 
          glimpse()
    ```


## Question 2

The `pokemon` data set contains information on (all?) Pokemon. Answer the following questions to help you practice your `dplyr` and `magrittr` skills.

- How many Pokemon are considered `Legendary`?

    ```{r}
    pokemon %>% 
      filter(Legendary == "True") %>% 
        summarise(n())   
    ```

- List the top five Pokeman, based on `Total`, whose `Type 1` is either `Grass` or `Fire`.

    ```{r}
    pokemon %>% 
      filter(`Type 1` == "Grass" | `Type 1` == "Fire" ) %>% 
      group_by(`Type 1`) %>% 
        top_n(5, Total)
    ```

- What are the mean and standard deviation of `HP` for each `Generation` of Pokemon?

    ```{r}
    pokemon %>% 
        group_by(Generation) %>% 
          summarize(myMean = mean(HP), mySTD = sd(HP))
    ```

- A Coefficient of Variation (CoV) is defined as the standard deviation divided by the mean ($\frac{s}{\bar{x}}$). Which `Generation` of Pokemon has the **lowest** Cov for `Attack`? 

    ```{r}
    pokemon %>% 
        group_by(Generation) %>% 
          summarize(CoV = sd(HP) / mean(HP)) %>% 
            arrange(CoV)
    ```

- Based on their `Type 2` characteristic, what are the Pokeman with the highest and lowest `Speed`?

    ```{r, eval = F}
    pokemon %>% 
      top_n(1, Speed)
      
      group_by(`Type 1`) %>% 
        #arrange(desc(Speed)) %>% 
          top_n(1, Speed) %>% 
            arrange(`Type 1`)
    ```
\ 


## Question 3

Import `uncSalaries.csv`, data on the salaries of the University of North Carolina's employees.
\ 

- What is the mean salary in the Neurosurgery department?

    ```{r}
    unc %>% 
      filter(dept == "Neurosurgery") %>% 
        summarise(meanSal = mean(totalsal, na.rm = T))
    ```
\ 

- Return a data frame with employee’s in the Neurosurgery department making more than $500,000. Why might these professors be so well paid?

    ```{r}
    unc %>% 
      filter(dept == "Neurosurgery", totalsal > 500000)
    ```
\ 

- What is the total amount that full time Dermatology employees get paid

    ```{r}
    unc %>% 
      filter(dept == "Dermatology", fte == 1.0) %>% 
        summarise(sum(totalsal))
    ```
\ 

- Create a data frame called `radio_dept` whose rows are the employees from the Radiology department.
    + include only the following columns: `name`, `position`, `age`, `nonstsal`, `totalsal`.
    + order the employees by salary
    
    ```{r}
    unc %>% 
      filter(dept == "Radiology") %>% 
        select(name, position, age, nonstsal, totalsal) %>% 
          arrange(desc(totalsal))
    ```
\ 

- Create a data frame called `dept_summary` whose rows are the departments and whose columns are: department size, mean department salary, median department salary, and maximum salary (using totalsal for salary).

    ```{r}
    dept_summary <- unc %>% 
      group_by(dept) %>% 
        summarise(deptSize = n(), medSal = median(totalsal, na.rm = T), maxSal = max(totalsal, na.rm = T))
    ```
\ 

- Order the departments by highest mean salary and print the 10 highest paid departments.
    ```{r}
    unc %>% 
      group_by(dept) %>% 
        summarise(meanSal = mean(totalsal, na.rm = T)) %>% 
          arrange(desc(meanSal)) %>% 
            top_n(10, meanSal)
    ```
\ 

- Order the departments by highest median salary and print the 10 highest paid departments.
    ```{r}
    unc %>% 
      group_by(dept) %>% 
        summarise(medSal = median(totalsal, na.rm = T)) %>% 
          arrange(desc(medSal)) %>% 
            top_n(10, medSal)
    ```

- Why do these lists differ? If you were asked for the top 10 best paid departments at UNC which summary would you choose and why?
