---
title: "Case Study in `ggplot2`"
author: "Paul Intrevado"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(fig.width = 5, fig.height = 3.5, fig.align = 'center') 
```

```{r libraries, include = F}
library(tidyverse)
library(hflights)
library(magrittr)
library(scales)
library(reshape2)
library(gcookbook)
```

## Question 1

The `reshape2` package contains a dataset named `tips`, which contains information on dining trasactions. Summary of the data follows:

```{r}
glimpse(tips)
```

Create the following graphs:

\ 

(a) Plot `time` on x-axis, total_bill on the y-axis, colored by `smoker` and shaped by `sex`.

\ 

```{r}
tips %>% 
  ggplot() +
    geom_point(aes(x = time, y = total_bill, shape = sex, colour = smoker))
```
\newpage
(b) `jitter` the previous plot so points are more visible

```{r}
tips %>% 
  ggplot() +
    geom_point(aes(x = time, y = total_bill, shape = sex, colour = smoker), position = "jitter")
```

\ 

(c) Create a graph with `time` on the x-axis, mean `tip` on the y-axis, shaped and coloured by `sex`.

```{r}
tips %>% 
  mutate(prctTip = tip / total_bill) %>% 
    group_by(sex, time) %>% 
      summarize(meanPrct = mean(prctTip, na.rm = T)) %>% 
        ggplot() +
          geom_line(aes(x = time, y = meanPrct, group = sex), linetype = "dotted", alpha = 0.6) +
          geom_point(aes(x = time, y = meanPrct, shape = sex, colour = sex), size = 6, alpha = 0.4) +
          xlab("\nMeal Time") +
          ylab("Mean Percent Tip\n") +
          scale_y_continuous(labels = percent, limits = c(0.15, 0.18), breaks = seq(0.15, 0.18, by = 0.005)) + 
          theme(axis.text.x = element_text(colour="grey20",size=14,angle=30,hjust=.5,vjust=.5),
            axis.text.y = element_text(colour="grey20", size=10),  
            axis.title.x = element_text(colour="grey20", size=14),
            axis.title.y = element_text(colour="grey20", size=14)) + 
            guides(colour = guide_legend("Gender"), shape = guide_legend("Gender")) +
            scale_colour_brewer(palette = "Set1")
```
\newpage

(d) Create a graph with `time` on the x-axis, `smoker` on the y-axis, colored by `sex` and sized by `prctTip`, with a minimalist `theme`.

```{r}
tips %>% 
  mutate(prctTip = tip / total_bill) %>% 
    group_by(sex, smoker, time) %>% 
      ggplot() +
      geom_point(aes(x = time, y = smoker, size = prctTip, color = sex), position = "jitter", alpha = 0.7) + 
      xlab("\nMeal Time") +
      ylab("Smoker/Non-Smoker\n") +
      theme(axis.text.x = element_text(colour="grey20",size=14),
            axis.text.y = element_text(colour="grey20",size=10),  
            axis.title.x = element_text(colour="grey20",size=14),
            axis.title.y = element_text(colour="grey20",size=14)) + 
            guides(colour = guide_legend("Gender"), size = guide_legend("% Tip")) +
            scale_colour_brewer(palette = "Set1") + theme_minimal()
```

\ 

(e) Create a stacked bar graph with `time` on x-axis, `count` on the y-axis, filled by `sex`.

```{r}
tips %>% 
  ggplot() +
    geom_bar(aes(x = time, fill = sex), alpha = 0.4) +
    scale_fill_discrete(name = "Gender")
```

\newpage
(f) Create a `dodge`d bar graph with `time` on x-axis, `count` on the y-axis, filled by `sex`.

```{r}
tips %>% 
  ggplot() +
    geom_bar(aes(x = time, fill = sex), position = "dodge", alpha = 0.5) +
    scale_fill_brewer(palette = "Set2", name = "Gender") 
```

(g) Create a bar graph with `time` on x-axis, `count` on the y-axis, and include the values of each bar.

```{r}
tips %>% 
  count(time) %>% 
  ggplot() +
  geom_bar(aes(x = time, y = n), stat = "identity", colour = "grey50") + 
  geom_text(aes(x = time, y = n, label = n), vjust = 1.5, colour = "white") +
  xlab("\nTime") +
  ylab("n\n")
```

\newpage

## Question 2

Create some fictional data using the following code:

```{r, echo = T}
set.seed(22)
x_int <- seq(10)
x_fact <- factor(seq(10))
y <- rnorm(10,2,13)
myDF <- data_frame(x_int, x_fact, y)
```

Note the `type`s of each of the columns in `myDF`.

(a) Create a line graph with `x_int` on the x-axis and `y` on the y-axis.

```{r}
myDF %>% 
  ggplot() +
    geom_line(aes(x = x_int, y = y))
```
\newpage

(b) Create a line graph with `x_fact` on the x-axis and `y` on the y-axis.

```{r}
myDF %>% 
  ggplot() +
    geom_line(aes(x = x_fact, y = y), group = 1)
```

\newpage

## Question 3

Create some fictional data using the following code:

```{r, echo = T}
set.seed(33)
day <- rep(seq(7), 2)
stockPrice <- rnorm(14, 2, 13)
company <- c(rep("GOOG", 7), rep("APPL", 7))
myDF <- data_frame(day, stockPrice, company)
```

\ 

(a) Create a line and point graph with `day` on the x-axis and `stockPrice` on the y-axis, grouped by `company`.

```{r}
myDF %>% 
  ggplot() +
  geom_line(aes(x = day, y = stockPrice, color = company)) +
  geom_point(aes(x = day, y = stockPrice, colour = company), shape = 21, fill = "white", size = 5) + 
  scale_color_brewer(palette = "Set1", name = "Company") +
  scale_y_continuous(labels =  dollar) +
  scale_x_continuous(breaks = seq(14)) + 
  xlab("\nWeek Day") +
  ylab("Stock Price\n")
```
\newpage

(b) Add a shaded blue bar to the previous graph, as shown.

```{r}
myDF %>% 
  ggplot() +
  annotate("rect", xmin = 2.5, xmax = 3.5, ymin = -20, ymax = 20, fill = "dodgerblue", alpha = 0.2) +
  geom_line(aes(x = day, y = stockPrice, color = company)) +
  geom_point(aes(x = day, y = stockPrice, colour = company), shape = 21, fill = "white", size = 5) + 
  scale_color_brewer(palette = "Set1", name = "Company") +
  scale_x_continuous(breaks = seq(14)) +
  xlab("\nWeek Day") +
  ylab("Stock Price\n")
```


\newpage

## Question 4

Create some fictional data using the following code:

```{r, echo = T}
set.seed(44)
year<- seq(1990, 2000)
numMovieReleases <- round(abs(rnorm(11, 2, 10)))
genre <- factor(sample(size = 11, c("Horror", "Drama", "SciFi"), replace = T))
myDF <- data_frame(year, numMovieReleases, genre)
```

Create an area graph with `year` on the x-axis and `numMovieRelases` on the y-axis.

```{r}
myDF %>% 
  ggplot() +
  geom_area(aes(x = year, y = numMovieReleases), group = 1, fill = "dodgerblue", alpha = 0.5, colour = "black") +
  scale_x_continuous(breaks = seq(1900, 2000)) +
  xlab("\nYear") +
  ylab("Number of Movie Releases\n") +
  theme_light()
```



\newpage

## Question 5

```{r, include = F}
myflights <- hflights
```

The `hflights` package contains a dataset named `hflights`, which provides information on `r format(nrow(myflights), big.mark = ",")` flights in 2011 leaving from Houston-based airports. How many flights departed per month? From IAH? From HOU? Create the following charts.

\ 

(a) Create a bar graph with `Origin` on the x-axis, `Frequency` on the y-axis, with `dodge`d bars for each month. 

```{r}
myflights %>% 
  count(Month, Origin) %>% 
    ggplot() + 
      geom_bar(aes(x = Origin, y = n, fill = factor(Month)), stat = "identity", position = "dodge") + 
      guides(fill = F) +
      scale_y_continuous(labels = comma) + 
      ylab("Frequency\n")
```

\newpage

(b) Create a bar graph with `Month` on the x-axis, `Frequency` on the y-axis, with `dodge`d bars for each `Origin`. 

```{r}
myflights %>% 
  count(Month, Origin) %>% 
    ggplot() + 
      geom_bar(aes(x = factor(Month), y = n, fill = Origin), stat = "identity", position = "dodge", alpha = 0.5) +
      xlab("\nMonth") +
      ylab("Frequency\n") + 
      scale_y_continuous(labels = comma) + 
      scale_fill_brewer(palette = "Set1")
```

\ 

(c) Generate a line graph with `Month` on the x-axis, `Frequency` on the y-axis, by `Origin`, and include a `dotted` horizontal line indicating the mean `Frequency` for each `Origin`, using a minimalist `theme`.

```{r, echo = F}
meanVals<- myflights %>% 
  count(Month, Origin) %>%
    group_by(Origin) %>% 
      summarize(meanVal = mean(n))


myflights %>% 
  count(Month, Origin) %>% 
    ggplot() + 
      geom_line(aes(x = Month, y = n, group = Origin, colour = Origin)) + 
      geom_point(aes(x = Month, y = n, colour = Origin), size = 4, alpha = 0.5 ) +
      geom_hline(yintercept = meanVals$meanVal[1], linetype = 'dotted') +
      geom_hline(yintercept = meanVals$meanVal[2], linetype = 'dotted') +
      scale_y_continuous(labels = comma) +    
      xlab("\nMonth") +
      ylab("Frequency\n") + 
      scale_colour_brewer(palette = "Set2") + 
      theme_minimal()
```




\newpage


## Question 6

```{r, include = F}
usPop <- uspopchange
```

The `gcookbook` package contains a dataset named `uspopchange`, which provides information on recent population changes in US States. We are interested in graphing the chnage in population across all states. Create the following graphs.

\ 

(a) Create a bar plot with `Percent Chnage` on the y-axis, `State` on the x-axis.

\ 

```{r}
usPop %>% 
  ggplot() +
    geom_bar(aes(x = State, y = Change/100), stat = "identity") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    ylab("% Change\n") + 
    scale_y_continuous(labels = percent)
```

\newpage

(b) Create a bar plot with `Percent Chnage` on the y-axis, `Region` on the x-axis, `dodge`d by state.

\ 

```{r}
usPop %>% 
  ggplot() +
    geom_bar(aes(x = Region, y = Change/100, fill = Abb), stat = 'identity',  position = "dodge") + 
    guides(fill = F) +
    ylab("% Change\n") +
    scale_y_continuous(labels = percent)
```

\ 

(c) Create a bar plot with `Percent Chnage` on the y-axis, `Region` on the x-axis, `dodge`d by state, arranged from lowest to highest within in `Region`.

\ 

```{r}
tempMap <-  usPop %>% 
  group_by(Region) %>% 
  arrange(Region, Change)

usPop$Abb <- factor(usPop$Abb, levels = tempMap$Abb, ordered = T)

usPop %>% 
  ggplot() +
  geom_bar(aes(x = Region, y = Change/100, fill = Abb), stat = 'identity',  position = "dodge") + 
  ylab("% Change\n") + 
  scale_y_continuous(labels = percent) + 
  guides(fill = F) 
```

\newpage

## Question 7

Import the data set `mlb_01.csv` from the course DropBox folder. 

(a) Generate a boxplot, with `Position.Name` on the x-axis, `Salary` in millions of dollars on the y-axis, and include, for each `Position.Name`, a marker indicating the mean `Salary`.

```{r, include = F}
mlbData <- read_csv("~/Desktop/mlb_01.csv")
```

```{r}
mlbMeans <- mlbData %>% 
              group_by(Position.Name) %>% 
                summarize(meanVals = mean(Salary))

mlbData %>% 
  ggplot() +
  geom_boxplot(aes(x = Position.Name, y = Salary/1000000), alpha = 0.5, colour = "grey50") +
  geom_point(data = mlbMeans, aes(x = Position.Name, y = meanVals/1000000), size = 5, shape = 10, colour = "green4", alpha = 0.5) + 
  xlab("\nPosition") +
  ylab("Salary ($M)\n")
```

(b) Generate a sactter plot, with `Position.Name` on the x-axis, `Salary` in millions of dollars on the y-axis, and include, red `labels` for all Japanese players and blue `labels` for all Dominican players.

```{r}
mlb_j <- mlbData %>% 
  filter(Japanese == 1)

mlb_d <- mlbData %>% 
  filter(Dominican == 1)

mlbData %>% 
  ggplot() +
  geom_point(aes(x = Position.Name, y = Salary/1000000), alpha = 0.5, colour = "grey50", size = 3) +
  geom_label(data = mlb_j, aes(x = Position.Name, y = Salary/1000000, label = Player), colour = "red", size = 2, fontface = "bold", alpha = 0.6, ) +
  geom_label(data = mlb_d, aes(x = Position.Name, y = Salary/1000000, label = Player), colour = "blue4", size = 2, fontface = "bold", alpha = 0.6, ) +
  xlab("\nPosition") +
  ylab("Salary ($M)\n")
```

\newpage


## Question 8

Generate fictional stock data including `year` and `stockPrice` using teh following code:

```{r, echo = T}
set.seed(99)
myDF <- data.frame(
  year = sample(1965:2015, 10000, replace = T), 
  stockPrice = rnorm(10000, 10, 2)
)
```


Generate a line graph with the mean `stockPrice` per `year` on the y-axis, the `year` on the x-axis, and include a `ribbon` that outlines the `min` and `max` `stockPrice` for a given year. 

```{r}
myDF %>% 
  group_by(year) %>% 
    summarise(
      s_mean = mean(stockPrice), 
      s_median = median(stockPrice), 
      s_min = min(stockPrice), 
      s_max = max(stockPrice)) %>% 
        ggplot() +
        geom_ribbon(aes(x = year, ymin = s_min, ymax = s_max), fill = "grey80", alpha = 0.7) +
        geom_line(aes(x = year, y = s_mean, group = 1), color = "navyblue") +  
        #geom_point(aes(x = year, y = s_mean), color = "navyblue", size = 1, alpha = 0.5) + 
        scale_y_continuous(labels = dollar, limits = c(0, 20),) +
        scale_x_continuous(limits = c(1965, 2015), breaks = seq(1965,2015, by = 5)) +
        xlab("\nYear") + ylab("Mean Yearly Stock Price\n")
```
