---
title: "Case Study in `dplyr`: `ggplot2` Movies"
author: "Paul Intrevado"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r libraries, include = F}
library(tidyverse)
library(ggplot2movies)
library(scales)
library(magrittr)
library(knitr)
myMovies <- movies
```

The `ggplot2movies` package contains a dataset named `movies`, which provides information on `r format(nrow(movies), big.mark = ",")` movies, datging as far back as `r min(movies$year)`! Here is a quick graphical summary of how many movies were realeased each year:

```{r}
myMovies %>% 
  group_by(year) %>% 
    count() %>% 
      ggplot(aes(x = year, y = n)) +
        geom_point() +
        geom_line() +
        xlab("\nYear") +
        ylab("Count\n") +
        scale_y_continuous(labels = comma) +
        scale_x_continuous(breaks = seq(1890, 2010, by = 10), limits = c(1890, 2010))
```
\ 

Aggregating by deacde, we observe that the numnber of movies produced by decade is monotonically non-decreaing (see Table below), save for the final deacde, perhaps because data was not collected fully through that decade.

```{r}
myMovies %<>%  mutate( 
decade = cut(year, breaks = seq(1890, 2010, by = 10), labels = paste(seq(1890, 2009, by = 10), "s", sep = ""), right = F, ordered_result = T))

myMovies %>% 
    group_by(decade) %>% 
      count %>% 
        knitr::kable(col.names = c("Decade", "Count"), format.args = list(big.mark = ","), align = c('c', 'c'))
```

Per decade, we can also track the total budget spent on movies.

```{r}
myMovies %>% 
  group_by(decade) %>% 
    select(decade, budget) %>% 
      summarise(totalBudget = sum(as.numeric(budget), na.rm = T)) %>% 
        knitr::kable(format.args = list(big.mark = ","), align = c('c', 'r'), col.names = c("Decade", "Total Budget"))
```

As judicious and transparent analysts, we must also disclose that the `budget` variable contains `r format(sum(is.na(myMovies$budget)), , big.mark = ",")`, the majority of which are in earlier decades. The following table offers a holistic view of the data we have.

```{r}
myMovies %>% 
  group_by(decade) %>% 
    select(decade, budget) %>% 
      summarise(
      totalBudget = sum(as.numeric(budget), na.rm = T), 
      totalNAs = sum(is.na(budget)), 
      totalMovies = n(),
      prctNAs = sum(is.na(budget)) / n() * 100
      ) %>%  
        knitr::kable(format.args = list(big.mark = ","), 
                     align = c('c', 'c', 'r', 'r', 'r'), 
                     col.names = c("Decade", "Total Budget", "Total NAs", "Total Movies", "% NAs"),
                     digits = 0)
```

It is very clear that the budget information we have is highly unrealiable as `r format(sum(is.na(myMovies$budget)) / nrow(myMovies) * 100, digits = 4)`% our of data are `NA`s. Looking at movie length, we observe that our data is far more reliable (see proceeding table).

```{r}
myMovies %>% 
  group_by(decade) %>% 
    select(decade, length) %>% 
      summarise(
      totalLenMin = sum(as.numeric(length), na.rm = T),
      meanLength = mean(as.numeric(length), na.rm = T),
      medianLength = median(as.numeric(length), na.rm = T),
      totalNAs = sum(is.na(length)), 
      totalMovies = n(),
      prctNAs = sum(is.na(length)) / n() * 100
      ) %>%  
        knitr::kable(format.args = list(big.mark = ","), 
                     align = rep('c', 7), 
                     col.names = c("Decade", "Total Movie Length (min)", "Mean Length", "Median Length", "Total NAs", "Total Movies", "% NAs"),
                     digits = 0)
```








