---
title: "Regulax Expressions Lab in `stringr`"
author: "Paul Intrevado"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(fig.width = 5, fig.height = 3.5, fig.align = 'center') 
```

```{r libraries, include = F}
library(tidyverse)
library(magrittr)
```

## Question 1

Using the `state.name` data, do the following:

  a) Find all states that contain at least one instance of he letter "z"
    ```{r}
    str_subset(state.name, "z")
    ```
  

  b) Find all states that contain two subsequent "s"'s, i.e.,"ss"
    ```{r}
    str_subset(state.name, "ss")
    ```

  c) Find all states that contain at least one but not more than two subsequent "s"'s, i.e.,"s" or "ss"
    ```{r}
    str_subset(state.name, "s{1,2}")
    ```
  d) Find all states that contain at lest two "i"'s, but the they need not be subsequent
    ```{r}
    str_subset(state.name, "i\\w*i")
    ```
  e) Find all states that have compound names, i.e., have two words separated by a space
    ```{r}
    str_subset(state.name, " ")
    ```

  f) Find all states that begin with either "North" or "South"
    ```{r}
    str_subset(state.name, "(North|South)")
    ```

\newpage

## Question 2

Using `fruit` data from stringr package, do the following:

  a) How many fruits are melons? 
    ```{r}
    str_subset(fruit, "melon")
    ```
  
  b) How many fruits are berries? 
    ```{r}
    str_subset(fruit, "berry")
    ```
  
  c) How many berries come from a single word versus a compound word?
    ```{r}
    length(str_subset(fruit, " berry"))
    length(str_subset(fruit, "berry"))
    ```
  
## Question 3
  
Locate 1st sequence of 1 or more consecutive numbers in the following character vector:

```{r}
x <- c("abcd", "a22bc1d", "ab3453cd46", "a1bc44d")
```

```{r}
str_locate(x, "\\d+")
```


## Question 4

  a) Write a regular expression that will match a typical US phone number.
  
    `"^//d{3}-//d{3}-//d{4}$"`
  
  b) What if that phone number begins with a "+1"? 
  
    `"^(\\+1)? ?//d{3}-//d{3}-//d{4}$"`
  
  c) What is the phone number lacks spaces?
  
    `"^(\\+1)? ?//d{3}-?//d{3}-?//d{4}$"`
  
  d) What if the area code is wrapped in round braces?
  
    `"^(\\+1)? ?(\\(?//d{3}\\)|//d{3})-?//d{3}-?//d{4}$"`
  
  e) What if, instead of dashes, someone decides to use spaces or a "."?
  
    `"^(\\+1)? ?(\\(?//d{3}\\)|//d{3})[-\\.]?//d{3}[-\\.]?//d{4}$"`

## Question 5

Write a regular expression that will match a gmail address.

  `"^\\w+@gmail.com$"`

