---
title: "Regulax Expressions Lab in `stringr`"
author: "Paul Intrevado"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(fig.width = 5, fig.height = 3.5, fig.align = 'center') 
```

```{r libraries, include = F}
library(tidyverse)
library(magrittr)
```

## Question 1

Using the `state.name` data, do the following:

  a) Find all states that contain at least one instance of he letter "z"

  b) Find all states that contain two subsequent "s"'s, i.e.,"ss"

  c) Find all states that contain at least one but not more than two subsequent "s"'s, i.e.,"s" or "ss"

  d) Find all states that contain at lest two "i"'s, but the they need not be subsequent

  e) Find all states that have compound names, i.e., have two words separated by a space

  f) Find all states that begin with either "North" or "South"

## Question 2

Using `fruit` data from stringr package, do the following:

  a) How many fruits are melons? 
  
  b) How many fruits are berries? 
  
  c) How many berries come from a single word versus a compound word?
  
## Question 3
  
Locate 1st sequence of 1 or more consecutive numbers in the following character vector:

```{r}
x <- c("abcd", "a22bc1d", "ab3453cd46", "a1bc44d")
```

## Question 4

  a) Write a regular expression that will match a typical US phone number.
  
  b) What if that phone number begins with a "+1"? 
  
  c) What is the phone number lacks spaces?
  
  d) What if the area code is wrapped in round braces?
  
  e) What if, instead of dashes, someone decides to use spaces or a "."?

## Question 5

Write a regular expression that will match a gmail address.
